---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
library(TEfits)
```

## Overview to Time-Evolving fits

The **TEfits** package streamlines nonlinear regression, such as that encountered in analysis of learning. The **TEfits** package is intended to assist in the implementation and interpretation of nonlinear regression with a heavy emphasis on interpretability of parameters. Parameters fit by **TEfits** are meant to reflect human-interpretable representations of time-evolving processes such as starting values, number of trials until 50% of change, or asymptotic values. Error functions, nonlinear ("change") functions linking parameters and time to predicted values, parameter and prediction boundaries, and goodness-of-fit indices are intended to be clear and adjustable. An equal emphasis is on ease of use: minimal arguments are necessary to begin using the primary function, `TEfit()`, and many common tasks are fully automated (e.g., optimization starting points, bootstrapping).

```{r simple_model, message=FALSE, warning=FALSE}

dat <- data.frame(resp=log(1:30),trialNum=1:30)
mod <- TEfit(dat[,c('resp','trialNum')])

plot(mod,plot_title='Time-evolving fit of artificial data')

summary(mod)
```

An example of a double-exponential fit using a Bernoulli response distribution, with 50 bootstrapped fits.

```{r model_a, message=FALSE, warning=FALSE}

dat <- data.frame(resp=log(1:30)/log(31),trialNum=1:30)
mod <- TEfit(dat[,c('resp','trialNum')], 
             errFun='bernoulli',
             changeFun='expo_double',
             bootPars=list(nBoots=50))

plot(mod,plot_title='Time-evolving fit of artificial data')

summary(mod)
```
