---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, include=FALSE}
library(TEfits)
```

# TEfits

[![DOI](https://zenodo.org/badge/225967950.svg)](https://zenodo.org/badge/latestdoi/225967950)
[![Build Status](https://travis-ci.com/akcochrane/TEfits.svg?branch=master)](https://travis-ci.com/akcochrane/TEfits)

## Overview to Time-Evolving fits

Behavioral data is described, interpreted, and tested using indices such as d prime, mean, or psychometric function threshold. The **TEfits** package serves to allow the same questions to be asked about time-evolving aspects of these indices, such as the starting level, the amount of time that the index takes to change, and the asymptotic level of that index. Nonlinear regression applied to time-evolving functions is made as intuitive and painless as is feasible, with many extensions if desired. 

The **TEfits** package has a heavy emphasis on interpretability of parameters. As far as possible, parameters fit by **TEfits** are meant to reflect human-interpretable representations of time-evolving processes. Error functions, nonlinear ("change") functions linking predicted values to parameters and time, parameter and prediction boundaries, and goodness-of-fit indices are intended to be clear and adjustable. An equal emphasis is on ease of use: minimal arguments are necessary to begin using the primary function, `TEfit()`, and many common tasks are fully automated (e.g., optimization starting points, bootstrapping).

## Simple model of exponential change

A basic model nonlinearly relating time to an outcome variable. The first argument is a data frame, with the first column being the response variable and the second column being the time variable.

```{r simple_model, message=FALSE, warning=FALSE}

# If necessary, install package:
# devtools::install_github('akcochrane/TEfits')

# generate artificial data:
dat_simple <- data.frame(response=log(2:31),trial_number=1:30)

# fit a `TEfit` model
mod_simple <- TEfit(dat_simple[,c('response','trial_number')])

plot(mod_simple,plot_title='Time-evolving fit of artificial data')

summary(mod_simple)
```

## Bootstrapped model with Bernoulli error function

An example of a learning fit using a Bernoulli response distribution, with 40 bootstrapped fits.

```{r model_boot, message=FALSE, warning=FALSE}
# generate artificial data:
dat <- data.frame(response=log(2:31)/log(32),trial_number=1:30)

# fit a `TEfit` model
mod_boot <- TEfit(dat[,c('response','trial_number')], 
             errFun='bernoulli',
             bootPars=tef_bootList(resamples = 40))
plot(mod_boot,plot_title='Time-evolving fit of artificial data with 95% CI from 40 bootstrapped fits')

summary(mod_boot)
```

## Fitting multiple models

An example of fitting a given model to subsets of data (e.g., individual participants within a behavioral study).

```{r model_groups, message=FALSE, warning=FALSE}
# generate artificial data:
dat <- data.frame(response=rep(dat$response,4)*seq(0,.2,length=120),trial_number=rep(1:30,4),group=rep(letters[1:4],each=30))

# fit a `TEfitAll` model
mod_4group <- TEfitAll(dat[,c('response','trial_number')], 
             groupingVar = dat$group,
             groupingVarName = 'Participant')

```

Note the warnings regarding rate parameters; identifiability is a major concern in nonlinear models, 
and `TEfits` attempts to notify the user of potentially problematic situations.

```{r plot_model_groups}

plot(mod_4group)

summary(mod_4group)
```

## Using a more common linear regression framework

In some cases (such as `mod_simple` above), similar performance can be attained using a nonlinear transformation of time as a predictor in a linear model. This method is plotted in green on top of the `mod_simple` results, with clearly near-identical fits.

```{r TElm}
# Fit a `lm` model, first computing the best nonlinear transformation for time:
mod_lm <- TElm(response~trial_number,dat_simple,timeVar = 'trial_number')

plot(mod_simple)

lines(dat_simple$trial_number,fitted(mod_lm),col='green',lty=2,lwd=2)

```

TElm parameter estimates:

`r  knitr::kable(round(data.frame(t(c(coef(mod_lm),rate=mod_lm$rate))),3))`

TEfit parameter estimates:

`r knitr::kable(round(data.frame(t(coef(mod_simple))),3))`

Note that `TEfit` provides start and asymptote parameters directly, while `TElm` provides 
start as an offset from asymptote (ie., `Intercept`).

# Performance disclaimer

**TEfits** comes with no guarantee of performance. Nonlinear regression can be very sensitive to small changes in parameterization, optimization starting values, etc. No universal out-of-the box implementation exists, and **TEfits** is simply an attempt to create an easy-to-use and robust framework for behavioral researchers to integrate the dimension of time into their analyses. **TEfits** may be unstable with poorly-behaved data, and using the option to bootstrap models is generally the best option for assessing the robustness of fits. In addition, running the same fitting code multiple times and comparing fit models should provide useful checks. All of these things take time, and **TEfits** is not built for speed; please be patient.

# Community guidelines

If you are having technical difficulties, if you would like to report a bug, or if you want to recommend features, it's best to open a Github Issue. Please feel welcome to fork the repository and submit a pull request as well.
