---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
library(TEfits)
```

## Overview to Time-Evolving fits

Data is described, interpreted, and tested using indices such as d prime, mean, or psychometric function threshold. The **TEfits** package serves to allow the same questions to be asked about time-evolving aspects of these indices, such as the starting level, the amount of time that the index takes to change, and the asymptotic level of that index. Nonlinear regression applied to time-evolving functions is made as intuitive and painless as is feasible, with many extensions if desired. 

The **TEfits** package has a heavy emphasis on interpretability of parameters. As far as possible, parameters fit by **TEfits** are meant to reflect human-interpretable representations of time-evolving processes. Error functions, nonlinear ("change") functions linking predicted values to parameters and time, parameter and prediction boundaries, and goodness-of-fit indices are intended to be clear and adjustable. An equal emphasis is on ease of use: minimal arguments are necessary to begin using the primary function, `TEfit()`, and many common tasks are fully automated (e.g., optimization starting points, bootstrapping).

```{r simple_model, message=FALSE, warning=FALSE}

dat <- data.frame(response=log(2:31),trial_number=1:30)
mod <- TEfit(dat[,c('response','trial_number')])

plot(mod,plot_title='Time-evolving fit of artificial data')

summary(mod)
```

An example of a learning fit using a Bernoulli response distribution, with 40 bootstrapped fits.

```{r model_boot, message=FALSE, warning=FALSE}

dat <- data.frame(response=log(2:31)/log(32),trial_number=1:30)
mod <- TEfit(dat[,c('response','trial_number')], 
             errFun='bernoulli',
             bootPars=list(nBoots=40))
plot(mod,plot_title='Time-evolving fit of artificial data with 95% CI from 40 bootstrapped fits')

summary(mod)
```

An example of fitting a given model to subsets of data (e.g., individual participants within a behavioral study).

```{r model_groups, message=FALSE, warning=FALSE}

dat <- data.frame(response=rep(dat$response,4)*seq(0,.2,length=120),trial_number=rep(1:30,4),group=rep(letters[1:4],each=30))

mod <- TEfitAll(dat[,c('response','trial_number')], 
             groupingVar = dat$group,
             groupingVarName = 'Participant')
plot(mod)

summary(mod)
```

Additional principles guiding the development of **TEfits**:

- Reliance only on base R (dependencies are few and optional: **psych**, **MASS**, **brms**)
- Good things come to those who wait: Speed is nice, but robustness is better.
